Intermediate representations are increasingly recognized as a key factor in the training dynamics and generation quality of Diffusion Transformers (DiTs). Existing self-guided methods such as layer synchronization and self-representation alignment improve these representations by directly aligning selected internal features, but they provide limited supervision on how hidden states evolve across depth. We propose depth shortcut prediction, a self-supervised objective that learns the cross-layer transition structure of a DiT. Given a source hidden state, source and target layer indices, and the diffusion timestep, a lightweight predictor estimates the target representation in direction--magnitude space by predicting both tokenwise directions and log-magnitude residuals. The predictor is trained with direct cross-layer supervision and bootstrap depth consistency, which encourages composed short transitions to match direct long transitions. Rather than treating cross-layer similarity as the objective itself, our method learns a conditional hidden-state transport model, providing dense supervision over the DiT depth trajectory and improving the structure of intermediate representations. The learned transport model further enables optional layer-skipping inference by reconstructing target hidden states before resuming the backbone. Experiments show that depth shortcut prediction improves representation learning and generative performance, with layer skipping emerging as an additional quality--speed tradeoff.
