\section{Introduction}

% Diffusion and flow-matching models are now a standard route to high-fidelity generation~\citep{ho2020ddpm,song2019score,song2021sde,lipman2022flow}. Transformer backbones such as DiT and SiT make these models scalable, but training remains expensive~\citep{peebles2023scalable,ma2024sit}. A useful way to improve training is to guide the hidden representations learned across transformer depth.

% Existing guidance is either external or internal. External methods align diffusion-transformer states with pretrained visual representations such as DINOv2 or vision-language features~\citep{yu2024repa,wang2025haste,leng2025repae,wang2026vaerepa,oquab2023dinov2,wang2024qwen2vl}. They provide strong targets, but add teacher computation and tie the recipe to a particular encoder. Internal methods avoid this dependency by using the diffusion transformer's own layer hierarchy~\citep{jiang2025sra,haghighi2026layersync}.

% The key obstacle is that each depth has its own representation coordinates. There often are semantic gap in representation quality across depth and often peaks before the final decoder-like blocks~\citep{mukhopadhyay2024,kim2025,stracke2025,haghighi2026layersync}. Direct internal alignment must therefore choose which weak and strong layers to match. This choice is heuristic, architecture-dependent, and sensitive to model depth; it also asks raw hidden states from different depths to meet despite representation mismatch.

% We propose \emph{ReSync}, a lightweight layer-aware transition predictor for flexible cross-depth internal alignment. Instead of aligning one manually selected layer pair, ReSync samples source--target layers and learns a shared predictor that maps an earlier hidden state into a later layer's representation space. Conditioning on the source layer, target layer, and denoising timestep lets the same module support arbitrary layer pairs without a separate adapter per pair.

% ReSync learns a bridge between layer distributions, so supervision follows the transition across depth instead of a few fixed endpoints. Uniform layer-pair sampling already provides a strong self-aligned signal; a pair-selection strategy then biases training toward transitions that are informative and learnable. This keeps internal alignment self-contained, less heuristic, and lightweight.

% Empirically, ReSync improves generation under matched training settings; on ImageNet $256\times256$ with SiT-B/2 and no classifier-free guidance, it improves FID from $25.78$ for the baseline and $21.32$ for LayerSync to $18.88$. Scaling results, objective ablations, pair-sampling diagnostics, and representation probes support the same message: a shared layer-aware transition predictor makes internal alignment more flexible while improving diffusion transformer representations.

% Our contributions are:
% \begin{itemize}[leftmargin=*,topsep=0.25em,itemsep=0.15em,parsep=0pt]
%     \item We introduce a lightweight layer-aware transition predictor for flexible cross-depth internal alignment in diffusion transformers.
%     \item We use stochastic layer-pair training and a simple pair-selection strategy to reduce reliance on hand-crafted layer choices.
%     \item We provide generation, scaling, ablation, representation, and overhead analyses showing that ReSync improves sample quality and internal representations with minimal standard-sampling cost.
% \end{itemize}


Diffusion and flow-matching models have become a leading paradigm for high-fidelity image generation~[CITE]. Recent Diffusion Transformer (DiT) architectures~[CITE] further improve scalability and generative quality by combining diffusion modeling with the strong sequence modeling capability of transformers. Despite these advances, training diffusion transformers remains computationally intensive, motivating growing interest in methods that improve representation learning to accelerate optimization and enhance generative performance~[CITE].

A prominent direction is representation alignment (RA) using semantically rich features extracted from external representation models (ERMs) such as DINOv2 or I-JEPA~[CITE]. While these approaches have demonstrated strong empirical performance, they introduce additional computational overhead and inherit limitations from the ERMs, including their inductive biases and representation quality. An alternative line of work instead leverages the hierarchical structure of diffusion transformers by aligning representations across different network depths~[CITE]. Compared with external representation alignment methods, these internal approaches are more self-contained and computationally efficient, but they still rely heavily on manually selected layer pairs for supervision.

However, internal representation alignment remains fundamentally challenging because activations at different transformer depths exhibit substantial semantic discrepancies. Shallow layers primarily mainly capture local patterns and low-level structural information, whereas deeper layers encode increasingly abstract semantic concepts~[CITE]. Consequently, directly matching hidden activations across depth is inherently difficult, causing existing methods to rely on heuristic, architecture-dependent layer selection strategies that often do not generalize across model scales or architectures. Moreover, even methods that introduce learnable transition predictors to bridge the semantic gap between layers~[] are generally designed for a specific source--target layer pair. Since the semantic relationship varies substantially across different layer combinations, predictors trained on one pair often transfer poorly to others.

To address these limitations, we propose \textbf{LARA}, a novel RA framework that enables flexible alignment across arbitrary layer pairs within a diffusion transformer. Our method is motivated by the observation that intermediate activations across network depth simultaneously contain shared structural regularities and layer-specific semantic variations. Building on this insight, LARA learns a lightweight layer-aware transition predictor that maps representations between different depths. The predictor captures common structural information through shared parameters while adapting to varying semantic gaps via conditioning on the source and target layer indices.

Unlike prior methods, LARA is both \emph{layer-agnostic} and \emph{architecture-agnostic}. By uniformly sampling arbitrary layer pairs during training, our framework removes the need for handcrafted layer selection and naturally supports alignment across the full transformer hierarchy. We further introduce advanced layer-pair sampling strategies, including [NAMES], which improve alignment quality while requiring minimal human intervention. Moreover, by distilling shared semantic structure across layers into a single compact predictor, LARA learns more robust and transferable representations that generalize effectively to downstream tasks.

We extensively evaluate LARA across multiple diffusion transformer architectures and benchmarks. Experimental results demonstrate consistent improvements in both generative quality and representation quality compared with prior methods, while introducing only minimal computational overhead. Comprehensive ablation studies further validate the effectiveness of layer-aware conditioning and shared transition modeling. Together, our results show that LARA provides an effective, scalable, and flexible framework for improving representation learning in diffusion transformers.