\section{Method}
\label{sec:method}

\subsection{Preliminaries}

We adopt the stochastic-interpolant formulation used by SiT~\citep{ma2024sit}, which provides a common view of flow-based and diffusion-based generative models. Given a data latent $x_0\sim p_{\mathrm{data}}$, Gaussian noise $\epsilon\sim\mathcal{N}(0,I)$, and timestep $t\in[0,1]$, the interpolation path is
\begin{equation}
x_t=\alpha_t x_0+\sigma_t\epsilon ,
\label{eq:stochastic_interpolant}
\end{equation}
where $\alpha_t$ and $\sigma_t$ control the mixture of data and noise. The schedule is chosen so that the path connects clean data to a tractable noise distribution, and generation follows the learned dynamics in the reverse direction. The instantaneous velocity of this path is
\[
u_t=\dot{\alpha}_t x_0+\dot{\sigma}_t\epsilon .
\]
A diffusion transformer $v_\zeta$ receives the noisy latent $x_t$, timestep $t$, and optional condition $c$, and is trained to predict this velocity by minimizing
\begin{equation}
\mathcal{L}_{\mathrm{FM}}
=
\mathbb{E}_{x_0,\epsilon,t,c}\left[
\left\|v_\zeta(x_t,t,c)-u_t\right\|_2^2
\right].
\label{eq:flow_matching_loss}
\end{equation}

\subsection{From Alignment to Transition Learning}

Prior work suggests that diffusion representations are highly non-uniform across depth: some intermediate blocks carry more useful semantic information, while the last blocks increasingly specialize toward decoding back to the latent space~\citep{mukhopadhyay2024,kim2025,stracke2025,haghighi2026layersync}. Existing self-contained methods use this structure by selecting internal activations and matching or regularizing them directly~\citep{jiang2025sra,haghighi2026layersync}. This gives useful guidance, but it still treats representations as static objects to be made similar.

ReSync keeps the same self-contained setting but changes the supervision from matching activations to learning transitions between them. The goal is not to make an early block reproduce a later block in the same raw feature space. Instead, we ask whether the earlier activation contains enough information for a small conditional predictor to recover the predictable part of the later activation. This gives earlier layers a directed learning signal while allowing different depths to preserve their own computational roles.

\subsection{ReSync Transition Prediction}

Let $A_\ell^t\in\mathbb{R}^{N\times C}$ denote the token activation after block $\ell$ when the model processes $(x_t,t,c)$. For a source layer $a$ and a target layer $b>a$, ReSync predicts the target activation from the source activation at the same denoising timestep:
\begin{equation}
T_{a\to b}^t(A)\equiv T_\psi(A,a,b,t),
\qquad
\hat A_{b\mid a}^t=T_{a\to b}^t(A_a^t).
\label{eq:resync_predictor}
\end{equation}
The predictor is shared across sampled layer pairs and conditioned on the denoising timestep. Thus, the projection-like module is no longer a local adapter for one selected pair, but an amortized transition model reused throughout the depth trajectory. Unless stated otherwise, $\mathbb{E}$ averages over data, noise, timesteps, and sampled layer indices; $d(\cdot,\cdot)$ is tokenwise cosine distance; and $\mathrm{sg}(\cdot)$ denotes stop-gradient. We train the predictor with
\begin{equation}
\mathcal{L}_{\mathrm{pred}}
=
\mathbb{E}\left[
d\left(\hat A_{b\mid a}^t,\mathrm{sg}(A_b^t)\right)
\right].
\label{eq:align_loss}
\end{equation}
The target activation is detached. This prevents the auxiliary loss from symmetrically pulling both layers together and makes the predictor responsible for bridging the source and target representations.

\subsection{Compositional Consistency}

The prediction loss alone can still behave like a collection of independent projection heads. To encourage a reusable internal transition rule, ReSync adds a compositional constraint. A long transition should agree with the composition of two shorter transitions:
\begin{equation}
\hat A_{c\mid a}^t
\approx
\hat A_{c\mid b\mid a}^t .
\label{eq:transition_composition}
\end{equation}
where $\hat A_{c\mid a}^t$ is the direct long transition and $\hat A_{c\mid b\mid a}^t$ is obtained by applying the transitions $a\to b$ and then $b\to c$. In the loss, we detach the source activation and use an EMA copy $\bar T$ to form a stable composed target. Let
\begin{equation}
\tilde A_a^t=\mathrm{sg}(A_a^t),
\qquad
\bar A_{b\mid a}^t=\bar T_{a\to b}^{\,t}(\tilde A_a^t),
\qquad
\bar A_{c\mid b\mid a}^t
=
\bar T_{b\to c}^{\,t}(\bar A_{b\mid a}^t).
\label{eq:composed_target}
\end{equation}
The consistency objective is then simply
\begin{equation}
\mathcal{L}_{\mathrm{cons}}
=
\mathbb{E}\left[
d\left(T_{a\to c}^t(\tilde A_a^t),\mathrm{sg}(\bar A_{c\mid b\mid a}^t)\right)
\right].
\label{eq:consistency_loss}
\end{equation}
This coupling is the main mechanism that distinguishes ReSync from pairwise self-alignment: the same predictor must explain local and longer-range changes along the depth trajectory.

\subsection{Transition Pair and Predictor Design}

The choice of transition pairs plays a role analogous to layer selection in direct self-alignment. ReSync uses a shared predictor, so we avoid designing a separate head for each source--target pair. Instead, we sample pairs from a bounded range of layer gaps. Very short gaps make the prediction task nearly trivial, while very long gaps can make the target representation too distant to provide stable supervision early in training. In practice, we use a mixture of short and medium gaps and ablate this sampling strategy in the experiments.

We also avoid relying on the final decoder-like blocks as the dominant targets. Prior analyses of diffusion representations indicate that semantic quality is heterogeneous across depth and that the last blocks become increasingly tied to output reconstruction~\citep{mukhopadhyay2024,kim2025,stracke2025,haghighi2026layersync}. We therefore treat the valid source range, target range, and gap distribution as predictor-design choices rather than fixed assumptions. Finally, the predictor itself is kept lightweight and shared across pairs through layer and timestep conditioning. This design keeps the method self-contained while testing whether one compact module can capture reusable internal transitions instead of memorizing pair-specific alignments.

\subsection{Objective and Extensions}

The main method uses same-timestep depth transitions because both source and target activations are available in one forward pass. More generally, the same predictor can be viewed as an intrinsic internal representation consistency function over the layer--time grid. In this extended form, it maps an activation from source layer $a$ and timestep $t$ to target layer $b$ and timestep $s$:
\begin{equation}
T_{a,t\to b,s}(A)\equiv T_\psi^{\mathrm{time}}(A,a,b,t,s),
\qquad
\hat A_b^s=T_{a,t\to b,s}(A_a^t).
\label{eq:cross_timestep_predictor}
\end{equation}
This layer--time extension requires target activations at another timestep $s$, so we report it only as a preliminary diagnostic rather than the main training recipe.

We also consider an optional output-level distillation term. Given a predicted activation, we pass it through the remaining backbone,
\begin{equation}
\hat y_{\mathrm{skip}}
=
F_{\zeta,b:L}(\hat A_b^t,t,c),
\qquad
\hat y_{\mathrm{full}}
=
F_{\zeta,b:L}(A_b^t,t,c),
\label{eq:skip_output}
\end{equation}
and match the resulting output to the full computation:
\begin{equation}
\mathcal{L}_{\mathrm{out}}
=
\left\|
\hat y_{\mathrm{skip}}
-
\mathrm{sg}(\hat y_{\mathrm{full}})
\right\|_2^2 .
\label{eq:output_distillation_loss}
\end{equation}
This term is not required for the main method. We use it only as an ablation to make the predictor more robust to its own errors and to encourage predicted activations to preserve information useful for the final velocity prediction.

We collect the auxiliary terms into a single predictor objective,
\begin{equation}
\mathcal{L}_{\mathrm{predictor}}
=
\mathcal{L}_{\mathrm{pred}}
+
\beta_{\mathrm{cons}}\mathcal{L}_{\mathrm{cons}}
+
\beta_{\mathrm{out}}\mathcal{L}_{\mathrm{out}} .
\label{eq:predictor_loss}
\end{equation}
The resulting training objective is
\begin{equation}
\mathcal{L}
=
\mathcal{L}_{\mathrm{FM}}
+
\lambda_{\mathrm{pred}}\mathcal{L}_{\mathrm{predictor}} .
\label{eq:total_loss}
\end{equation}
In all main experiments, $\beta_{\mathrm{out}}=0$, so ReSync is purely a training-time representation regularizer and adds no inference-time cost.
