\section{Experiments}
\label{sec:experiments}

This section evaluates ReSync as an amortized self-alignment mechanism: if useful internal guidance is better exposed through a learned interface, then a shared predictor should improve generation while producing measurable transition-level diagnostics. We organize the experiments around the following questions:
\begin{itemize}[leftmargin=*]
    \item How does each design choice and component of ReSync influence performance? (\Cref{sec:ablation_study})
    \item Does ReSync improve generation quality and training efficiency across domains and model sizes? (\Cref{sec:image_generation,sec:human_motion_generation})
    \item Does ReSync provide stronger representation guidance than baseline backbones and competing self-alignment methods? (\Cref{sec:comparison_representation_guidance,sec:representation_analysis})
    \item Does the learned predictor remain internally consistent and continue to produce reliable transitions on deeper, harder-to-predict layer pairs? (\Cref{sec:transition_diagnostics,sec:skip_prediction})
\end{itemize}

\subsection{Image Generation}
\label{sec:image_generation}
\label{sec:experimental_setup}

\paragraph{Implementation details.}
Unless otherwise stated, we follow the class-conditional ImageNet training setup of SiT~\citep{ma2024sit}. We train on ImageNet~\citep{deng2009imagenet} at $256\times256$ resolution using the preprocessing protocol of ADM~\citep{dhariwal2021diffusion}. Each image is encoded by the Stable Diffusion VAE from LDM~\citep{rombach2022ldm} into a latent tensor $z\in\mathbb{R}^{32\times32\times4}$. We evaluate the \texttt{B/2}, \texttt{L/2}, and \texttt{XL/2} SiT configurations with patch size $2$. ReSync trains the transition predictor jointly with the backbone using \Cref{eq:total_loss}. Unless specified, we use AdamW with learning rate $10^{-4}$ and batch size $256$. Additional details on architecture, transition-pair sampling, EMA decay, loss weights, training length, hardware, and compute budget are organized in \Cref{app:experimental_details}.

\paragraph{Evaluation.}
We report Frechet Inception Distance (FID)~\citep{heusel2017gans}, Inception Score (IS)~\citep{salimans2016improved}, precision, and recall~\citep{kynkaanniemi2019precision} using $50\mathrm{K}$ generated samples. We use the same reference statistics and evaluation implementation for all ReSync variants and baselines. Definitions for the diagnostic metrics that are not standard generation benchmarks are provided in \Cref{app:metric_definitions}.

\paragraph{Comparison methods.}
Our main comparisons are against the base SiT/DiT backbone and recent representation-regularized training methods. For self-contained guidance, we compare with Dispersive Loss~\citep{wang2025dispersive}, SRA~\citep{jiang2025sra}, and LayerSync~\citep{haghighi2026layersync}. For externally guided training, we compare with REPA~\citep{yu2024repa} and REED~\citep{wang2025flexible}. For completeness, we also report representative image-generation baselines: pixel-space diffusion models including ADM~\citep{dhariwal2021diffusion}, VDM++~\citep{kingma2023understanding}, Simple Diffusion~\citep{hoogeboom2023simple}, and CDM~\citep{ho2022cascaded}; latent diffusion and transformer-based models including LDM~\citep{rombach2022ldm}, U-ViT-H/2~\citep{bao2023all}, DiffiT~\citep{hatamizadeh2024diffit}, MDTv2-XL/2~\citep{gao2023mdtv2}, MaskDiT~\citep{zheng2023fast}, SD-DiT~\citep{zhu2024sddit}, DiT~\citep{peebles2023scalable}, and SiT~\citep{ma2024sit}; and autoregressive or predictive-generation baselines such as VAR~\citep{tian2024visual} and D-JEPA~\citep{chen2024jepa}. Detailed baseline descriptions and the parameter-matching protocol are deferred to \Cref{app:baseline_descriptions,app:baseline_protocol}.

\paragraph{Results.}
\Cref{tab:imagenet_no_cfg} reports the main no-CFG ImageNet $256\times256$ comparison under matched SiT training settings, while \Cref{tab:imagenet_cfg} reports a system-level comparison with CFG. We reserve the numerical entries for final runs and reproduced baselines; the result summary should state both final sample quality and training-budget efficiency once the table is filled. \textbf{[PLACEHOLDER: summarize no-CFG FID gains, CFG system-level comparison, and whether ReSync narrows the gap to externally guided methods such as REPA/REED]}.

\begin{table}[!htbp]
    \centering
    \begin{minipage}[t]{0.45\textwidth}
        \captionsetup{font=small,justification=raggedright,singlelinecheck=false}
        \captionof{table}{\textbf{FID comparisons of class-conditional generation on ImageNet $256\times256$.} No classifier-free guidance (CFG)~\citep{ho2022classifier} is used. Numerical entries are placeholders.}
        \label{tab:imagenet_no_cfg}
        \vspace{0.35em}
        \scriptsize
        \setlength{\tabcolsep}{2.1pt}
        \renewcommand{\arraystretch}{0.78}
        \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lccc@{}}
            \toprule
            Model & \#Params & Epochs & FID $\downarrow$ \\
            \midrule
            \texttt{SiT-B/2} & 130M & 80 & \textbf{25.78} \\
            \quad + Dispersive & 130M & 80 & \textbf{25.23} \\
            \quad + SRA & 130M & 80 & \textbf{25.08} \\
            \quad + LayerSync & 130M & 80 & \textbf{21.32} \\
            \quad + ReSync & 130M & 80 & \textbf{18.88} \\
            \midrule
            \texttt{SiT-L/2} & 458M & 80 & \textbf{[TBD]} \\
            \quad + SRA & 458M & 80 & \textbf{[TBD]} \\
            \quad + LayerSync & 458M & 80 & \textbf{[TBD]} \\
            \quad + ReSync & 458M & 80 & \textbf{[TBD]} \\
            \midrule
            \texttt{SiT-XL/2} & 675M & 80 & \textbf{[TBD]} \\
            \quad + SRA & 675M & 80 & \textbf{[TBD]} \\
            \quad + LayerSync & 675M & 80 & \textbf{[TBD]} \\
            \quad + ReSync & 675M & 80 & \textbf{[TBD]} \\
            \midrule
            \texttt{SiT-XL/2} & 675M & 200 & \textbf{[TBD]} \\
            \quad + SRA & 675M & 200 & \textbf{[TBD]} \\
            \quad + LayerSync & 675M & 200 & \textbf{[TBD]} \\
            \quad + ReSync & 675M & 200 & \textbf{[TBD]} \\
            \midrule
            \texttt{SiT-XL/2} & 675M & 400 & \textbf{[TBD]} \\
            \quad + SRA & 675M & 400 & \textbf{[TBD]} \\
            \quad + LayerSync & 675M & 400 & \textbf{[TBD]} \\
            \quad + ReSync & 675M & 400 & \textbf{[TBD]} \\
            \bottomrule
        \end{tabular*}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{0.51\textwidth}
        \captionsetup{font=small,justification=raggedright,singlelinecheck=false}
        \captionof{table}{\textbf{System-level comparison on ImageNet $256\times256$ with CFG~\citep{ho2022classifier}.} Numerical entries are placeholders.}
        \label{tab:imagenet_cfg}
        \vspace{0.35em}
        \scriptsize
        \setlength{\tabcolsep}{1.8pt}
        \renewcommand{\arraystretch}{0.80}
        \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lccccc@{}}
            \toprule
            Model & Epochs & FID$\downarrow$ & IS$\uparrow$ & Pre.$\uparrow$ & Rec.$\uparrow$ \\
            \midrule
            \multicolumn{6}{@{}l}{\textit{Pixel diffusion}} \\
            ADM-U & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            VDM++ & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & -- & -- \\
            Simple diffusion & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & -- & -- \\
            CDM & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & -- & -- \\
            \midrule
            \multicolumn{6}{@{}l}{\textit{Autoregressive image generation}} \\
            VAR-d20 & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            D-JEPA-L & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            \midrule
            \multicolumn{6}{@{}l}{\textit{Latent diffusion, U-Net}} \\
            LDM-4 & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            \midrule
            \multicolumn{6}{@{}l}{\textit{Latent diffusion, Transformer + U-Net hybrid}} \\
            U-ViT-H/2 & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            \midrule
            \multicolumn{6}{@{}l}{\textit{Latent diffusion, Transformer}} \\
            MaskDiT & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            SD-DiT & \textbf{[TBD]} & \textbf{[TBD]} & -- & -- & -- \\
            DiT-XL/2 & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            SiT-XL/2 & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            \quad + REPA & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            \quad + REED & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            \quad + Dispersive & \textbf{[TBD]} & \textbf{[TBD]} & -- & -- & -- \\
            \quad + LayerSync & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            \quad + ReSync & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
            \bottomrule
        \end{tabular*}
    \end{minipage}
\end{table}

\subsection{Text-Conditional Human Motion Generation}
\label{sec:human_motion_generation}

To test whether ReSync transfers beyond image generation, we additionally evaluate text-conditioned human motion generation on HumanML3D~\citep{guo2022generating}. Given a text description, the model generates a sequence of human poses represented as 3D joint trajectories.

\paragraph{Implementation details.}
We follow the MDM setup~\citep{tevet2022mdm} and use a transformer backbone with the same training budget for the baseline and ReSync variant. The main comparison trains MDM with and without ReSync for $600\mathrm{K}$ iterations. \textbf{[PLACEHOLDER: specify exact backbone depth, selected transition layers, predictor size, loss weights, batch size, and evaluation checkpoint]}.

\paragraph{Evaluation metrics.}
We report FID and R-Precision (top 3) using the standard HumanML3D evaluation protocol. Definitions and reporting conventions are provided in \Cref{app:metric_definitions}. Lower FID is better, while higher R-Precision indicates better text-motion retrieval alignment.

\paragraph{Results.}
\Cref{tab:humanml3d_results} reports the HumanML3D comparison. The result summary should emphasize whether the same signal-extracting transition objective improves a smaller, non-image transformer without changing the generative task. \textbf{[PLACEHOLDER: summarize FID and R-Precision improvements once results are available]}.

\begin{table}[!htbp]
    \centering
    \caption{\textbf{Quantitative results for text-conditional human motion generation on HumanML3D.} Numerical entries are placeholders until final evaluation.}
    \label{tab:humanml3d_results}
    \small
    \setlength{\tabcolsep}{8pt}
    \renewcommand{\arraystretch}{0.9}
    \begin{tabular}{@{}lccc@{}}
        \toprule
        Method & Iter. & FID $\downarrow$ & R-Precision $\uparrow$ \\
        \midrule
        MDM & 600K & \textbf{[TBD]} & \textbf{[TBD]} \\
        + ReSync (Ours) & 600K & \textbf{[TBD]} & \textbf{[TBD]} \\
        \bottomrule
    \end{tabular}
\end{table}

\FloatBarrier

\subsection{Ablation Study}
\label{sec:ablation_study}

We first isolate the contribution of each ReSync design choice. The ablation starts from the same backbone trained without ReSync, then adds layer-conditioned prediction, timestep conditioning, compositional consistency, output consistency, and the weak diversity term. We report the no-diversity variant separately to verify that the self-alignment interface itself improves convergence and sample quality, rather than attributing the gains solely to cosine-squared repulsion. We then compare direct and projected self-alignment controls to test whether projection exposes useful layer signals rather than merely adding adapter capacity. Detailed layer-pair, timestep, and predictor-capacity sweeps are moved to \Cref{app:resync_ablation_protocols}.

\begin{table}[!htbp]
    \centering
    \begin{minipage}[t]{0.48\textwidth}
        \captionsetup{font=small,justification=raggedright,singlelinecheck=false}
        \captionof{table}{\textbf{Ablation of ReSync components.} Lower FID is better; higher IS is better.}
        \label{tab:resync_ablation}
        \vspace{0.25em}
        \scriptsize
        \setlength{\tabcolsep}{2.5pt}
        \renewcommand{\arraystretch}{0.88}
        \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcccccc@{}}
            \toprule
            Variant & Trans. & Time & Comp. & Extra & FID $\downarrow$ & IS $\uparrow$ \\
            \midrule
            Baseline & -- & -- & -- & -- & \textbf{[TBD]} & \textbf{[TBD]} \\
            + transition & \checkmark & -- & -- & -- & \textbf{[TBD]} & \textbf{[TBD]} \\
            + time cond. & \checkmark & \checkmark & -- & -- & \textbf{[TBD]} & \textbf{[TBD]} \\
            + composition & \checkmark & \checkmark & \checkmark & -- & \textbf{[TBD]} & \textbf{[TBD]} \\
            + output cons. & \checkmark & \checkmark & \checkmark & out & \textbf{[TBD]} & \textbf{[TBD]} \\
            Full ReSync & \checkmark & \checkmark & \checkmark & out+div & \textbf{[TBD]} & \textbf{[TBD]} \\
            \bottomrule
        \end{tabular*}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{0.48\textwidth}
        \captionsetup{font=small,justification=raggedright,singlelinecheck=false}
        \captionof{table}{\textbf{Projection interfaces for self-alignment.} These controls test whether projection improves direct internal alignment, especially for weak source layers.}
        \label{tab:projection_transition_ablation}
        \vspace{0.25em}
        \scriptsize
        \setlength{\tabcolsep}{1.5pt}
        \renewcommand{\arraystretch}{0.88}
        \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcccc@{}}
            \toprule
            Method & Source & Proj. & FID $\downarrow$ & CKA $\uparrow$ \\
            \midrule
            Baseline & -- & -- & \textbf{[TBD]} & -- \\
            LayerSync-style & selected & -- & \textbf{[TBD]} & \textbf{[TBD]} \\
            + projector & selected & \checkmark & \textbf{[TBD]} & \textbf{[TBD]} \\
            Weak-layer align. & weak & -- & \textbf{[TBD]} & \textbf{[TBD]} \\
            Weak-layer + proj. & weak & \checkmark & \textbf{[TBD]} & \textbf{[TBD]} \\
            All-layer + proj. & all & \checkmark & \textbf{[TBD]} & \textbf{[TBD]} \\
            \bottomrule
        \end{tabular*}
    \end{minipage}
\end{table}

\FloatBarrier

\subsection{Transition Diagnostics}
\label{sec:transition_diagnostics}

Generation metrics alone cannot show whether ReSync is using the intended mechanism. We therefore measure whether the learned predictor behaves like a reusable interface for internal guidance rather than a fixed-pair adapter. The main diagnostics are transition prediction error on sampled layer pairs, held-out prediction error on unseen layer pairs, compositional error between direct and composed transitions, and the relationship between these errors and generation quality. These measurements do not prove a causal mechanism, but they test whether the empirical gains are consistent with transition-mediated self-alignment.

\begin{table}[!htbp]
    \centering
    \caption{\textbf{Transition-level diagnostics.} Placeholder table for testing whether ReSync learns a reusable internal-guidance interface. Lower prediction and compositional errors are better; stronger correlation with generation metrics would support the interpretation that transition quality is related to sample quality.}
    \label{tab:transition_diagnostics}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{lccccc}
        \toprule
        Method variant & Train error $\downarrow$ & Held-out layer $\downarrow$ & Comp. error $\downarrow$ & FID $\downarrow$ & Error--FID corr. \\
        \midrule
        Pair-specific projector & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        Shared transition without consistency & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        ReSync shared transition & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        \bottomrule
    \end{tabular}}
\end{table}

\FloatBarrier

\subsection{Comparison with Representation-Guided Training}
\label{sec:comparison_representation_guidance}

ReSync is designed to provide representation guidance without an external encoder or an external representation task. We therefore compare against methods that use pretrained visual representations, self-guided internal alignment, or other representation regularizers. This comparison should report both final sample quality and training efficiency, since external encoders can improve convergence while adding additional training-time computation.

\begin{table}[!htbp]
    \centering
    \caption{\textbf{Comparison with representation-guided diffusion training.} Placeholder table for methods using external encoders, external representation tasks, or internal self-guidance.}
    \label{tab:comparison_external_guidance}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{lccccc}
        \toprule
        Method & External encoder & External task & FID $\downarrow$ & IS $\uparrow$ & Training cost \\
        \midrule
        Baseline backbone & -- & -- & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        REPA-style alignment & \checkmark & -- & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        SRA / self-alignment & -- & -- & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        LayerSync-style alignment & -- & -- & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        ReSync & -- & -- & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        \bottomrule
    \end{tabular}}
\end{table}

\FloatBarrier

\subsection{Skipped-Depth Prediction}
\label{sec:skip_prediction}

Although ReSync is trained as a self-alignment regularizer, its predicted states can also be tested during inference. We evaluate skipped-depth prediction by replacing a block interval with the predicted hidden state and then resuming the remaining backbone. The primary measurements are output error against the full model, generation quality under skipped execution, and wall-clock speed. We treat this as an auxiliary capability rather than the central claim.

\begin{table}[!htbp]
    \centering
    \caption{\textbf{Skipped-depth prediction.} Placeholder table for evaluating whether predicted hidden states can support optional layer skipping.}
    \label{tab:skip_prediction}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{lccccc}
        \toprule
        Variant & Skipped blocks & Output error $\downarrow$ & FID $\downarrow$ & Speedup $\uparrow$ & Notes \\
        \midrule
        Full backbone & 0 & -- & \textbf{[TBD]} & $1.0\times$ & -- \\
        ReSync without output consistency & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        ReSync with output consistency & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} & \textbf{[TBD]} \\
        \bottomrule
    \end{tabular}}
\end{table}

\FloatBarrier

\subsection{Representation Analysis}
\label{sec:representation_analysis}

Finally, we test whether ReSync changes the backbone's intermediate representations in ways that are consistent with the proposed mechanism. The main diagnostics follow \Cref{fig:sit_xl_layer_statistics}: activation RMS or token norm measures statistical mismatch across layers, while CKA measures representational mismatch. We treat spatial-layout checks, affinity-based similarity, cosine similarity, SVCCA, rank spectra, temporal CKA heatmaps, and gradient-alignment diagnostics as supplementary diagnostics in \Cref{app:layer_statistics_projection}; measurement definitions and reporting conventions are given in \Cref{app:metric_definitions}.

\FloatBarrier
