\section{Limitations and Future Work}

Our current study focuses on validating ReSync under the experimental settings available within our compute budget: \textbf{[PLACEHOLDER: completed datasets/modalities]}. Due to the cost of large-scale generative pretraining, we do not yet evaluate ReSync on open-domain text-to-video pretraining or broad multimodal training at the same scale as recent frontier models. The final scope of completed runs, compute budget, and unfinished experiments should be reported explicitly: \textbf{[PLACEHOLDER: final compute budget and unfinished runs]}. Accordingly, any claims about video or large-scale multimodal generation should be viewed as future directions rather than established empirical conclusions in this submission.

We believe the extension to video is conceptually natural because ReSync is self-contained: it uses the diffusion transformer's own internal depth trajectory rather than relying on a strong external representation encoder. This property is particularly relevant for domains where pretrained encoders may be weaker, less standardized, or mismatched to the target generative distribution. However, this hypothesis requires direct validation on video data, including comparisons against external-representation alignment methods and self-contained alignment baselines.

ReSync also suggests a possible path toward layer-skipping inference through output-consistent transition prediction. In this paper, we treat this as an auxiliary ablation rather than the main claim, because a complete acceleration claim would require measuring generation quality, wall-clock speed, memory, and error accumulation under different skip lengths, predictor sizes, and output-consistency weights: \textbf{[PLACEHOLDER: skip-inference ablation]}. Future work should further study when predicted intermediate states remain stable under long skips and whether transition consistency reduces sensitivity to layer-pair selection.

Another future direction is extending ReSync from same-timestep depth transitions to transitions over both layer and denoising time. Such a formulation would require target activations from additional timesteps and a separate evaluation of training cost, stability, and inference utility, so we leave it outside the main method.

\section{Conclusion}

We introduced ReSync, a self-contained objective for learning compositional internal representation transitions in Diffusion Transformers. Rather than directly aligning two hidden states, ReSync predicts the predictable component of a later-layer representation from an earlier-layer representation at the same denoising stage, and enforces consistency between composed and direct depth transitions. This framing combines the simplicity of internal self-alignment with an amortized transition model for training-time representation regularization. Cross-timestep prediction and layer-skipping inference remain future directions rather than primary claims.
